---
title: "California Coast Cetacean Species Richness"
output:
  distill::distill_article:
    code_folding: true
    css: postcards.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(raster)
library(tidyverse)
library(here)
library(sf)
library(fasterize)
library(rnaturalearth)
library(rnaturalearthhires)
library(rgeos)
```

## Summary
This is a map showing the species richness of cetaceans along the California Coast. I read in a raster stack of cetacean species ranges. I considered a species present if its probability of presence was greater than or equal to 60%. I then counted the number of species present within each raster cell and produced the map showing species richness. The raster was paired with a map of California from the `rnaturalearth` package to help visualize the data and understand the location of it.

## Data Wrangling
```{r}
map <- ne_states(country = "united states of america", returnclass = "sf") %>% 
  dplyr::select(postal) %>% 
  filter(postal == "CA")
```

## Creating the Species Richness Raster
```{r, fig.cap = "**Figure 1:** The species richness, meaning the number of species present, of cetaceans on the California Coast. The presence of the cetaceans was based on a threshold of at least a probability of occurence of 0.6 for each species. Data from: Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016). AquaMaps: Predicted range maps for aquatic species. www.aquamaps.org."}
# read in the data
my_files <- list.files(path = here("cetacean_data", "ca_cetaceans"), full.names = TRUE)

species_stack <- raster::stack(my_files)

species_df <- raster::rasterToPoints(species_stack) %>%
  as.data.frame() %>% 
  pivot_longer(cols = Balaenoptera_acutorostrata:Ziphius_cavirostris, names_to = "species", values_to = "probability") %>% 
  mutate(presence = case_when(
    probability >= 0.6 ~ 1,
    probability < 0.6 ~ 0
  )) %>% 
  drop_na() %>% 
  filter(presence == 1) %>% 
  group_by(x,y) %>% 
  count(presence)

map <- st_transform(map, st_crs(species_stack))

ggplot() +
  geom_raster() +
  geom_raster(data = species_df, aes(x = x, y = y, fill = n)) +
  geom_sf(data = map, aes(inherit.aes = TRUE)) +
  coord_sf(xlim = c(-125, -114), ylim = c(32, 42)) +
  scale_fill_continuous(low = "lightblue", high = "blue", name = "Number of Cetacean Species") +
  labs(title = "Species Richness of Cetaceans on the California Coast",
       x = "Longitude",
       y = "Latitude")

```