---
title: "Text Analysis"
output:
  distill::distill_article:
    code_folding: true
    css: postcards.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(tidytext)
library(textdata)
library(ggwordcloud)
library(here)
library(patchwork)
library(plotly)
library(scales)
```

## Summary
This is a text analysis that compares the top words used in Matthew to those in 1st Corinthians. There is also a AFINN sentiment analysis of all of the words in both books. To conduct this analysis, the text files first needed to be cleaned up and separated into one word per row. 

## Data Wrangling
```{r}
# my own stop words data frame
word <- c("thou", "thee", "ye", "thy", "hath", "cock", "ass", "yea", "saith", "hast", "shalt")
stop_words_special <- data.frame(word)

# read in and wrangle matthew
matthew <- read.delim(here("bible_text", "bible", "matthew.txt")) %>% 
  separate(col = ST..MATTHEW, into = c("ch", "verse", "text", "extra", "extra2"), sep = ":") %>% 
  unite(col = text, text:extra:extra2, sep = " ") %>% 
  mutate(text = str_replace(text, pattern ="NA NA", replacement = " ")) %>% 
  mutate(text = str_replace(text, pattern ="NA", replacement = " ")) %>% 
  mutate(text = str_trim(text)) %>% 
  mutate(text = str_squish(text)) %>% 
  mutate(ch = str_squish(ch)) %>% 
  mutate(verse = str_squish(verse)) %>% 
  mutate(text = str_to_lower(text)) %>% 
  unnest_tokens(word, text) %>% 
  anti_join(stop_words) %>% 
  anti_join(stop_words_special)

# read in and wrangle first corinthians
first_corinth <- read.delim(here("bible_text", "bible", "1corinth.txt")) %>% 
  separate(col = X1.CORINTHIANS, into = c("ch", "verse", "text", "extra", "extra2"), sep = ":") %>% 
  unite(col = text, text:extra:extra2, sep = " ") %>% 
  mutate(text = str_replace(text, pattern ="NA NA", replacement = " ")) %>% 
  mutate(text = str_replace(text, pattern ="NA", replacement = " ")) %>% 
  mutate(text = str_trim(text)) %>% 
  mutate(text = str_squish(text)) %>% 
  mutate(ch = str_squish(ch)) %>% 
  mutate(verse = str_squish(verse)) %>%
  mutate(text = str_to_lower(text)) %>% 
  unnest_tokens(word, text) %>% 
  anti_join(stop_words) %>% 
  anti_join(stop_words_special)

# counts
matthew_counts <- matthew %>% 
  count(word)

first_corinth_counts <- first_corinth %>% 
  count(word)

# top 5
top_5_matt <- matthew_counts %>%
  arrange(-n) %>% 
  slice(1:5)

top_5_first_corinth <- first_corinth_counts %>% 
  arrange(-n) %>% 
  slice(1:5)

# Visualizing the counts
matthew_graph <- ggplot(data = top_5_matt, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "lightblue") +
  theme_minimal() +
  coord_flip() +
  labs(x = "Word",
       y = "Count",
       title = "Top 5 Words in the Gospel of Matthew")

first_corinth_graph <- ggplot(data = top_5_first_corinth, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "lightblue4")+
  theme_minimal() +
  coord_flip() +
  labs(x = "Word",
       y = "Count",
       title = "Top 5 Words in First Corinthians")
```

## Compound Graph
```{r, fig.cap = "The top five words that appear in the Gospel of Matthew and in First Corinthians. Words like 'the' and 'are' have been removed from the analysis. Since this is from the New King James Version, other words like 'thou', 'thee', 'ye', 'thy', 'hath', 'yea', 'saith', 'hast', 'shalt'."}
# compound graph
graph_compound <- (matthew_graph/first_corinth_graph) & plot_layout(guides = "collect")

graph_compound
```

## Matthew Top 100 Word Cloud
```{r, fig.cap = "A word cloud of the top 100 words from the Gospel of Matthew, New King James Version."}
# word cloud
matthew_cloud_counts <- matthew_counts %>% 
  arrange(-n) %>% 
  slice(1:100)

matthew_cloud <- ggplot(data = matthew_cloud_counts, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, size = n)) +
  scale_size_area(max_size = 15) +
  labs(title = "Top 100 Words in the Gospel of Matthew")

matthew_cloud
```


## First Corinthians Top 100 Word Cloud
```{r, fig.cap = "A word cloud of the top 100 words from First Corinthians, New King James Version."}
first_corinth_cloud_counts <- first_corinth_counts %>% 
  arrange(-n) %>% 
  slice(1:100)

first_corinth_cloud <- ggplot(data = first_corinth_cloud_counts, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, size = n)) +
  scale_size_area(max_size = 15) +
  labs(title = "Top 100 Words in First Corinthians")

first_corinth_cloud
```

## AFINN Analysis
```{r, fig.cap = "Affin sentiment analysis, where positive numbers show positive sentiment and negative numbers indicate negative sentiment. The words 'cock' (rooster) and 'ass' (donkey) have been removed because their meaning is misconstrued by the sentiment analysis."}
matthew_afinn <- matthew %>% 
  inner_join(get_sentiments("afinn"))

afinn_matt_means <- matthew_afinn %>% 
  group_by(ch) %>% 
  summarize(mean_afinn = mean(value)) %>% 
  mutate(ch = as.numeric(ch)) %>% 
  arrange(ch)

affin_matt_graph <- ggplot(data = afinn_matt_means,
       aes(x = ch, y = mean_afinn)) +
  geom_col(fill = "lightblue") +
  coord_flip() +
  theme_minimal() +
  labs(y = "Mean Afinn Sentiment",
       x = "Chapter",
       title = "Sentiment of the Gospel of Matthew")

first_corinth_afinn <- first_corinth %>% 
  inner_join(get_sentiments("afinn"))

afinn_corinth_means <- first_corinth_afinn %>% 
  group_by(ch) %>% 
  summarize(mean_afinn = mean(value)) %>% 
  mutate(ch = as.numeric(ch)) %>% 
  arrange(ch)

affin_corinth_graph <- ggplot(data = afinn_corinth_means,
       aes(x = ch, y = mean_afinn)) +
  geom_col(fill = "lightblue4") +
  coord_flip() +
  theme_minimal() +
  labs(y = "Mean Afinn Sentiment",
       x = "Chapter",
       title = "Sentiment of First Corinthians")

affin_compound <- (affin_matt_graph/affin_corinth_graph) & plot_layout(guides = "collect")

affin_compound
```

**Data Citation:** Stewarton Bible School. Stewarton, Scotland. March 2000. Holy Bible: King James Version. http://www.stewartonbibleschool.org/bible/text/index.html
